### 2020-01-21 paper/paper.tex
More Paper Updates
Fleshed out the writing some, added real figures where there were fruit bats before, added captions

HOGG: THIS IS NOW ALL WRONG... We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
the two-dimensional $(x,y)$-position on the detector and the
wavelength $\lambda$
\begin{equation}
\lambda(x,y,m,n) = f(x,y,m;\theta_{n})
\quad ,
\end{equation}
where $\theta_{n}$ is a big blob of parameters for this exposure.

@@

If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so
small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could
be replaced with a tiny model
\begin{equation}
\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
\quad ,
\end{equation}
where
$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the
spectrograph,
the $a_{nk}$ are scalar amplitudes,
and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the
fiducial calibration.
The challenge is to learn these basis functions from the data and get
the $K$ amplitudes $a_{nk}$ for every exposure $n$.

HOGG: Say that there are many ways to do this, but we use PCA for now.
Note that PCA is correct in the limit that everything is very high in SNR.

\subsection{Data} \label{sec:data}
We present results using data from \expres\, the EXtreme PRecison Spectrograph.  \expres\ is an envionmentally stabilized, fiber-fed, $R=137,000$, optical spectrograph (\lz{CJ and RB's papers}).  \expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g. \lz{Wilken+ 2012, Molaro+ 2013, Probst+ 2014, from RB's paper}).  ThAr exposures are taken at the beginning and end of each night.  LFC exposures are interspersed between science exposures every ~15 minutes.  
(Discussion of number of orders covered or number of lines in LFC and/or ThAr?)

\lz{Discussion of line fitting?  Though the current line fitting causes me recurring nightmares.  I should really take the time to look into at least the peak fitting more.}

For the results presented in this paper, we work with the 1D, optimally extracted data.  This removes the y-dependence from equations 1 and 2.

\expres\ exposures are separated into different ``instrumental epochs," which correspond to changes to the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources (\lz{(worth including table that explicitly defines epochs?)}).  Because significant instrumental changes break the assumption of only low-dimensional variations in the spectrograph hardware, we treat each instrument epoch independently.

\subsection{Hierarchical De-Noising of Calibration Frames} \label{sec:denoising}
Let $\{exp_n\}$ be the set of calibration exposures within an instrument epoch used to construct $g_k(x,y,m)$.  We first identify the full list of lines we can expect to find in a calibration image from this epoch.  A line is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength of that line, $\lambda$.  Let $\{l_{m,\lambda}\}$ be the complete list of unique lines found in $\{exp_n\}$.

For each line, $l_{m,\lambda}$, we then find the fitted line center, $x_{n,l}$, in pixel space, for that line for each exposure.  If this line is missing from an exposure (e.g. it was covered by noise or off the edge of its usual order), we insert a $NaN$ for that line center instead.

Lines that are missing from greater than some percent of exposures are cut.  Exposures that are missing greater than some percent of lines are also cut (\lz{This will miss low signal exposures that suffer most in the first few orders but can recover in later orders; can't decide if this is a feature or a bug}).  This threshold percentage is 50\% for LFC exposures and 30\% for ThAr exposures.

Missing line center measurements are replaced with denoised estimates using iterative PCA.  For each missing line center measurement, we initialzie their value to the mean of the nearest $n$ exposures in time, where $n$ is typically $9$ exposuresr.  We then find a fiducial calibration of the spectrograph,  $g_0(x,y,m)$, for which we use the mean line center value for each identified line.  We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and fiducial line center for each exposure.  This PCA gives us $a_{nk}$ and $g_k(x,y,m)$.

Using those values with equation 2 and some small integer (e.g. 2) for K, we can reconstruct a de-noised matrix of line centers for each exposure.  Missing line center measurements are then replaced with values from this reconstruction.  We do this iteratively until the values of the missing line centers change by less than 0.01\% of a pixel.  This convergence suggests that the PCA is no longer being swayed by outliers and is truly ``de-noised."

\begin{algorithm}
\SetAlgoLined
\textbf{Inputs}: \;
\While{change in bad pixels $>$ 0.01\%}{
	$g_0(x,y,m) = \overline{\{x_{n,l}\}}$\;
	find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x_{l,n}-g_0(x,y,m))$\;
	let $a_{n,k} = U\cdot \Sigma$ and $V = g_k(x,y,m)$\;
	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ where $K=2$\;
	$\{x_{n,l}\} = \lambda(x,y,m,n)$ where $\{x_{n,l}\}$ is $NaN$
	}
\caption{Hierarchical De-Noising}
\end{algorithm}

\subsection{Non-Parametric Calibration} \label{sec:nonparam}
We can interpolate $a_{nk}$ with respect to time to determine $\lambda(x,y,m,n)$ at any time.  Once we have the line center for each line, which has an associated, known wavelength, we can interpolate the known wavelengths over line centers to any position in x in a given order.  For example, interpolating onto every integer x will generate wavelengths for each pixel in an order.

We implement with respect to time using a linear spline, which is sufficient to capture the linear variations we see over a night (see \ref{fig:nightlyVariation}).  Interpolation of wavelengths over live centers is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.  This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an order while also enforcing monotonicity.

\begin{figure}[h]
\centering
\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
\caption{Show amplitudes of principal components w.r.t. time to show  that a linear interpolation is reasonable.  Should be a close up of one night or multiple nights (b/c the larger structure does not make it seem like a linear interpolation is a good idea.}
\label{fig:nightlyVariation}
\end{figure}

@@

\subsection{Choice Choices} \label{sec:choices}
(Can use results of training/validation or cross-tests to argue for different tests?  Or should we go all the way through to RVs.

\subsection{Value of K}
\begin{figure}[h]
\centering
\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
\caption{SVD eigenvectos showing sharp drop off after 2 (may be different if bluer orders are included}
\label{fig:eigenvalues}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
\caption{eigenvectors through the sixth one showing mostly noise following the first few}
\label{fig:eigenvectors}
\end{figure}

\subsection{Interpolation in x Choice}
Reason for not linear interpolation:
Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels is greater at greater wavelengths.  This means that the function of wavelength given pixel across an order will necessarily be concave down everywhere.  Consequently, linear interpolation of this function will systematically give erroneously low values.

\begin{figure}[h]
\centering
\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
\caption{We could have a figure that compares linear interpolation, cubic interpolation, and monotonic interpolation along with a histogram of the residuals}
\label{fig:xinterpolation}
\end{figure}

Allowing for a non-zero smoothing parameter is bad as it allows for smoothness we have no reason to believe is true of the wavelength solution.  However, restricting to a monotonic wavelength solution follows from how the optics work since we are finding the solution order by order.

\subsection{Order of Interpolation}
Now doing time then x direction.  No real reason for the switch except this makes more sense (well, to me) and is faster.

\section{Results} \label{sec:results}
\subsection{Self Tests} \label{sec:test-self}
First, let's have a sanity check.  Using a linear interpolator with no smoothing, we predict wavelengths for the denoised line center values, $\lambda(x,y,m,n)$, for an exposure used in the training set.  This returns identical values to the assigned true wavelength for each line within machine error for both ThArs and LFCs (see Fig. \ref{fig:sanity test}).

\begin{figure}[h]
\centering
\includegraphics[width=0.48\textwidth]{Figures/lfcLfcDenoised.pdf}
\includegraphics[width=0.48\textwidth]{Figures/tharTharDenoised.pdf}
\caption{results of interpolating to denoised value (probably not necessary, more an internal check of whether the code is at a base level behaving how we would expect.}
\label{fig:sanityTest}
\end{figure}

We then predict wavelengths for the measured line centers for an exposure that was used to construct the basis.  The linear interpolation in time should return identical PCA coefficients for this time.  Differences in wavelengths will arise from a combination of 1) noise in the measured line centers and 2) any interpolation error of lines to other x positions.  Figure \ref{fig:selfTest} shows a histogram of the residuals.

\begin{figure}[h]
\centering
\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
\caption{include median/spread of residuals?}
\label{fig:selfTest}
\end{figure}

\subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
A more realistic test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis and weights with PCA.  We can then use the results from that PCA to predict wavelengths for lines in the calibrations images that were left out.  We will call the used calibration images the ``training set," and the left out exposures the ``validation set."

We chose to use 90\% of available exposures to train the PCA, and leave the remaining 10\% for validation.  Exposures were sorted into either set at random.  This will test how well PCA coefficients are interpolated to other times outside of the training set and therefore how well line centers can be reconstructed at all times.  As before, differences in wavelength will also be drawn from how well line centers are measured and error in interpolating to those measured values.  This second source of error was independently quantified in the second test described in the previous section.

\begin{figure}[h]
\centering
\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
\caption{results from training and validation super-imposed over self test probably}
\label{fig:tNvTest}
\end{figure}

\subsection{Cross Tests} \label{sec:test-cross}
Similar to the training and validation tests, we can also use one calibration source to construct the PCA basis and weights and the other to test the results.  For example, we can use LFC exposures to construct the wavelength solution, and then use it to predict wavelengths for measured line centers of ThAr lines, which will have their own assigned true wavelengths.  Likewise, we can use ThAr exposures to predict the wavelengths of LFC lines.

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{Figures/lfcgood.pdf}
\caption{}
\label{fig:crossTest}
\end{figure}

\section{Comparisons to Other Methods} \label{sec:comparisons}
I think we probably have to use the even/odd method to compare.

\subsection{Comparison to Parametric Model}
Previously, the \expres\ pipeline the $\theta_{n}$ from equation 1 comprises the
amplitudes of some 9th-ish-order polynomial in $x$ and $m$ for each $n$.
If so, they have
\begin{equation}
\lambda(x,y,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
\quad ,
\end{equation}
where the $c_{nij}$ are coefficients unique to exposure $n$ but smoothed by a third order polynomial over time.
Further, they were finding the coefficients $c_{nij}$ that
minimize an objective $Q$ that is something like the L2-norm:
\begin{equation}
Q = ||\lambda(x,y,m,n) - \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j||_2^2
\quad .
\end{equation}

Even in this simple context, where every calibration image $n$ is
treated as its own unique flower, there are improvements to be
made.
For one, the sums should not be from 0 to 9, but instead
\begin{equation}
\sum_{i=0}^9\sum_{j=0}^{9-i}
\quad ,
\end{equation}
because that is the definition of 9th order.  We find that fitting to 8th order is sufficient, faster, and leaves less residual pattern than the original 9th degree fit.
For another, the objective function could be made soft to permit
catastrophic outliers without destroying the fit. (\lz{we did not do this})
We might recommend the iteratively reweighted least squares (IRLS).
This would make the fitting more robust.
For yet another, there are rescaling issues for the products $x^i\,m^j$ to
protect the fitting from near-singularities or bad conditioning of the
linear-algebra operators.

To test each method, we took an LFC exposure and separated the even and odd lines.  We then used the even lines to predict wavelengths for the odd lines and vice versa.  For example, for the parametric model, we fit a 2D, 8th-degree polynomial to the odd lines.  We then evaluated this polynomial at the positions of the even lines to assign them wavelengths.  We then reversed the process by fitting a polynomial to the even lines and evaluating that polynomial at the positions of the odd lines.  Differences between the polynomial assigned wavelengths and the assigned wavelengths of each line are shown in Figure \ref{fig:dsnMFit}.    For the non-parametric model, cubic spline interpolation was used in place of a polynomial fit.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/polyval2d.pdf}
\caption{\lz{Unclear whether it is worth a direct comparison with the old EXPRES pipeline} Residuals of a parametric model.  A 2D, 8th-degree polynomial fit was used to predict the wavelengths of each line.   Left shows the difference between the predicted and true wavelength for each line (color) across each order.  Right is a histogram of these residuals.  Residuals have been translated to velocity shifts.}
\label{fig:polyValFit}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/designMatrix.pdf}
\caption{Residuals of a parametric model fit using a design matrix.  A 2D, 8th-degree polynomial fit was used to predict the wavelengths of each line.  Otherwise, similar to Figure \ref{fig:polyValFit}.}
\label{fig:dsnMFit}
\end{figure}

\subsection{Comparison to Non-Parametric, Non-Hierarchical Model}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/noHierc.pdf}
\caption{Residuals of a non-parametric model.  A PCHIP cubic spline interpolator was used to predict the wavelengths of each line.  otherwise similar to Figure \ref{fig:dsnMFit}.}
\label{fig:intpFit}
\end{figure}

>>>>

We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
the two-dimensional $(x,y)$-position on the detector and the
wavelength $\lambda$
\begin{equation}
\lambda(x,y,m,n) = f(x,y,m;\theta_{n})
\quad ,
\end{equation}
where $\theta_{n}$ is a big blob of parameters for this exposure.

@@

If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could be replaced with a tiny model
\begin{equation}
\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
\quad ,
\end{equation}
where
$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the spectrograph,
the $a_{nk}$ are scalar amplitudes,
and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the fiducial calibration.

The challenge is to learn these basis functions from the data and get the $K$ amplitudes $a_{nk}$ for every exposure $n$.  There are many ways to discern these basis functions.  Here, we present a model using principal component analysis (PCA).  This is the correct thing to do in the limit where we have very high SNR, as is undeniably the case with calibration images.

\subsection{Data} \label{sec:data}
We present results using data from \expres\, the EXtreme PRecison Spectrograph.  \expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph (\lz{CJ and RB's papers}).  \expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g. \lz{Wilken+ 2012, Molaro+ 2013, Probst+ 2014, from RB's paper}).  ThAr exposures are taken at the beginning and end of each night.  LFC exposures are interspersed between science exposures every ~15 minutes.

The method and results discussed here are based off of 1260 LFC exposures and ThAr exposures over 90 days.  Exposures were taken whenever EXPRES was assigned time and so are randomly distriubted throughout the nine days

\lz{(Discussion of number of orders covered or number of lines in LFC and/or ThAr?)}

For the results presented in this paper, we work with the 1D, optimally extracted data.  This removes the y-dependence from equations 1 and 2.  Gaussian profiles are fit to each line in the 1D, extracted data.  The mean of each fitted Gaussian is taken to be the center of the line.  This process is described in detail in \lz{RP Paper} and considered out of scope of this paper.  The method described proceeds from a given list of orders, line wavelengths, and line centers for each exposure.

\expres\ exposures are separated into different ``instrumental epochs," which correspond to atypical, large instrument changes that affected the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources (\lz{can cite either Ryan's paper for more detail?}).  Because significant instrumental changes break the assumption of only low-dimensional variations in the spectrograph hardware, we treat each instrument epoch independently.

\subsection{Hierarchical De-Noising of Calibration Frames} \label{sec:denoising}
Let $\{exp_n\}$ be the set of calibration exposures within an instrument epoch.  These are the exposures to be used in constructing $g_k(x,y,m)$.  We first identify the full list of lines we can expect to find in a calibration image from this epoch.  A line is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength of that line, $\lambda$.  We define $\{l_{m,\lambda}\}$ to be the complete list of unique lines found in $\{exp_n\}$.

For each line, $l_{m,\lambda}$, we then find the fitted line center, $x_{n,l}$, in pixel space, for that line in each calibration exposure.  If this line is missing from an exposure (e.g. it was covered by noise or off the edge of its usual order), we insert a $NaN$ for that line center instead.

Lines that are missing from greater than some percent of exposures are completely cut from the analysis.  Exposures that are missing greater than some percent of lines are also cut.  We use a threshold percentage of 50\% for both lines and exposures.  This largely affects the first lines on either end of an order and exposures with spuriously low signal.  Typically, about 0.2\% of lines and cut for LFC based calibration and 0.25\% of files are cut.

Missing line center measurements are replaced with denoised estimates using iterative PCA.  For each missing line center measurement, we initialize their value to the mean of the nearest $n$ exposures in time, where $n$ is typically $9$ exposures.  We then find a fiducial calibration of the spectrograph,  $g_0(x,y,m)$, for which we use the mean line center value for each identified line.  We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and mean line center for each exposure.  This process gives us $a_{nk}$ and $g_k(x,y,m)$.

Using those values with equation 2 and some small integer (e.g. 2) for $K$, we can reconstruct a de-noised matrix of line centers for each exposure.  Missing line center measurements are replaced with values from this reconstruction.  We iterate until the values of the missing line centers change by less than 0.01\% of a pixel.  This convergence suggests that the PCA is no longer being swayed by outliers and is truly ``de-noised."

\begin{algorithm}
\SetAlgoLined
\textbf{Inputs}: ${m,\lambda}$, $\{l_{m,\lambda}\}$ \;
\While{change in bad lines $>$ 0.01\%}{
	$g_0(x,y,m) = \overline{\{x_{n,l}\}}$\;
	find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x_{l,n}-g_0(x,y,m))$\;
	let $a_{n,k} = U\cdot \Sigma$ and $V = g_k(x,y,m)$\;
	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ where $K=2$\;
	$\{x_{n,l}\} = \lambda(x,y,m,n)$ where $\{x_{n,l}\}$ was initially $NaN$
	}
\caption{Hierarchical De-Noising}
\end{algorithm}

Following the de-noising process, we check for deviant line centers.  This is aimed at catching lines that have been mis-identified or poorly fit, resulting in erroneous line centers.  Outliers are identified via a simple sigma cut for each line.  Any line center that is reported at more than $3\sigma$ away from the mean line center value is relabeled a ``bad line,'' and we repeat the process of hierarchical de-noising. (\lz{Should probably add something about why the sigma cut works})

\subsection{Non-Parametric Calibration} \label{sec:nonparam}
The process of hierarchical de-noising gives us $a_{nk}$, which captures the variability of the instrument.  We can interpolate $a_{nk}$ with respect to housekeeping data to determine $\lambda(x,y,m,n)$ at any time.  In this paper, we show results where $a_{nk}$ is interpolated with respect to time, but it is possible to incorporate other information such as temperature, pressure, or telescope position. (\lz{Does this belong in the discussion?})

Once we have the line center for every line in $\{l_{m,\lambda}\}$, each of which is associated with a known wavelength, we can interpolate the known wavelengths over line centers to any position in x in a given order.  For example, interpolating onto every integer x will generate wavelengths for each pixel in an order.

@@

\subsection{Choice Choices} \label{sec:choices}
Implementing \name required choosing some global variables and methods that we believe is optimal for constructing a high-fidelity wavelength solution.  We discuss these choices here and how we came to each decision.

\subsection{Value of K}
The value of $K$ determines how many principal components are used to reconstruct de-noised line centers.  This $K$ needs to be large enough that all variability in the spectrograph is captured.  Too large, and the reconstruction will begin to incorporate noise, defeating the purposes.

We settled on a $K$ value of 2.  Figure \ref{fig:eigenvectors} shows the eigenvectors of the PCA plotted on a log scale for both a LFC-based calibration scheme (blue), and a ThAr-based one (green).  In both cases, there is a precipitous drop off at one, indicating that the later principal components have a much smaller contribution.

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{Figures/ss.pdf}
\caption{Eigenvectors from PCA as step functions.  The results for using only LFC exposures (blue) and only using ThAr exposures (green) are both shown on a log scale.  After 1, there is a order of magnitude drop for both the LFC and the ThAr.}
\label{fig:eigenvectors}
\end{figure}

Figures \ref{fig:pcLfc} and \ref{fig:pcThar} shows the first 6 principal components for the LFC and ThAr respectively.  In both cases there is clear structure in the first and second principal components.  It is doubly encouraging that this structure appears the same in both the case with the LFC and the ThAr.

The later principal components show mostly noise or are swayed by large noise sources.  For example, the band on the left side of each order in principal component 3 for the LFCs is likely due to lower signal on the edges of the order from the blaze function, giving more uncertain line center measurements.  The diagonal banding visible in the sixth principal component can be traced to what we believe are detector defects.

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{Figures/pcsLfc6.png}
%\includegraphics[width=\textwidth]{Figures/pcsLfc6.pdf}
\caption{The first six principal components using LFC exposures.  There is clear structure for the first two principal components.  \lz{Principal components 3-5 show mostly noise (do they though?)}.  Principal component 6 shows structure from what we believe to be detector defects.}
\label{fig:pcLfc}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics[width=\textwidth]{Figures/pcsThar6.png}
%\includegraphics[width=\textwidth]{Figures/pcsThar6.pdf}
\caption{The first six principal components using ThAr exposures.  Note the similarities with the LFC result shown in Figure \ref{fig:pcLfc}.}
\label{fig:pcThar}
\end{figure}

\subsection{Interpolation in Time}
Figure \ref{fig:nightlyVariation} shows the amplitude of the first and second principal component with respect to time on the left.  Again, it is comoforting to note that the LFC and ThAr versions show the same shape.  Though there are no large trends over the ninety days of data shown here, there is a clear linear trend within each night (see Fig. \ref{fig:nightlyVariation} right).  We therefore interpolate these amplitudes with respect to time using a simple linear interpolation scheme.

\lz{linear interpolation actually doesn't look super great for the second principal component}

\begin{figure}[h!]
\centering
\includegraphics[width=0.56\textwidth]{Figures/pcAs.pdf}
\includegraphics[width=0.42\textwidth]{Figures/pcAs_byDay.pdf}
\caption{Amplitude of the first two principal components show as a function of time (left) or fraction of a day (right).  Points mark the times at which we have calibration data.  Lines show the result of a linear interpolation.  Left: amplitudes as a function of time for both LFC (blue) and ThAr (green).  The top plot shows the amplitudes for the first prinicpal component while the bottom plot shows the amplitudes for the second principal component.  Note that the shape of the curves is similar regardless of calibration source.  Recall from Figures \ref{fig:pcLfc} and \ref{fig:pcThar} that the corresponding principal components are also fairly similar.
Right: amplitudes of the first two principal components determined by the LFC exposures only.  Here exposures are grouped by day and plotted with respect to what time of day the exposure was taken.  A linear interpolation through all the LFC exposures taken in one day are shown as black line.  Points are colored by the MJD of the exposure.}
\label{fig:nightlyVariation}
\end{figure}

\subsection{Interpolation in Pixel}
Interpolation of wavelengths over line centers is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.  This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an order while also enforcing monotonicity.

Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels is greater at greater wavelengths.  This means that the function of wavelength given pixel across an order will necessarily be concave down everywhere.  Consequently, linear interpolation of this function will systematically give erroneously low values.

A more classic cubic spline interpolator runs into issues with lines in the ThAr that are blended and therefore very close together.  Very close lines cause huge deviations from the correct wavelengths, as seen in Figure \ref{fig:xinterp}, which highlights deviations caused by a blend in a specific order of ThAr exposures.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/intpx_tests.pdf}
\caption{Results from different interpolation schemes in the pixel direction.
Left: resultant wavelength solution curves when using ThAr lines (green points) to predict wavelengths for LFC lines (black points).  The green points are being used to predict values for the black points.  Different interpolation schemes are shown in different colors.  The largest deviation comes with the classic cubic spline (pink).  The other models are hard to distinguish.
Right: histograms of residuals for each method.  The linear interpolation model (yellow) is asymmetric with disproportionately positive residuals.  The classic cubic spline residuals (pink) has very spread out residuals, the extremes of which have been cut off in the figure.  The residuals from an interpolation scheme that incorporates smoothing (blue) are non-Gaussian and rather plateau about zero.  Residuals from the PCHIP interpolator (orange) has the strongest peak at 0 and is roughly Gaussian.}
\label{fig:xinterp}
\end{figure}

These huge digressions can be avoided by allowing for some smoothing in the interpolation.  In Figure \ref{fig:xinterp}, we show an example in blue using scipy's Univarate Spline method.  While the result appears to follow the calibration lines much better, the smoothing ultimately causes larger residuals that are spatially correlated (Fig. \ref{fig:xinterp}, right).

For all orders, the edges will be overestimated while the middle will be underestimated.  (\lz{Do I need to include a figure for this?  Or should I just not go into this much detail})  This indicates that the smoothing is causing the resultant wavelength solution to underestimate the curvature of the pixel-wavelength relation.  It also reduces the benefit we get from moving to a non-parametric model by enforcing a smoothness we have no reason to believe is true in the data.

We instead turn to the PCHIP interpolator, which damps down huge deviations in the traditional cubic spline by requiring the resulting interpolated function to be monotonic.  This is a constraint we know is true from the optics since interpolation in the pixel direction is carried out order-by-order.  From Figure \ref{fig:xinterp}, it is clear that the PCHIP interpolator returns the lowest residuals.

\subsection{Order of Interpolation}
Now doing time then x direction.  \lz{No real reason for the switch except this makes more sense (well, to me) and is faster.}

\section{Results} \label{sec:results}
\subsection{Self Tests} \label{sec:test-self}
First, let's have a sanity check.  (\lz{Is it worth mentioning this at all?}) Using a linear interpolator with no smoothing, we predict wavelengths for the denoised line center values, $\lambda(x,y,m,n)$, for an exposure used in the training set.  This returns identical values to the assigned true wavelength for each line within machine error for both ThArs and LFCs (see Fig. \ref{fig:sanity test}).

\begin{figure}[h]
\centering
\includegraphics[width=0.48\textwidth]{Figures/lfcLfcDenoised.pdf}
\includegraphics[width=0.48\textwidth]{Figures/tharTharDenoised.pdf}
\caption{Residuals for predicted wavelengths for the denoised x-values of an exposure used to train the PCA.   As expected, the residuals are zero within machine error for the case where LFCs are used and where ThArs are used.}
\label{fig:sanityTest}
\end{figure}

We then predict wavelengths for the \textit{measured} line centers for an exposure that was used to construct the basis.  The linear interpolation in time should return identical PCA coefficients for this time.  Differences in wavelengths will arise from a combination of 1) noise in the measured line centers and 2) any interpolation error of lines to other x positions.  Figure \ref{fig:testHists} shows a histogram of the residuals for both the LFC (left) and the ThAr (right) in black.

\subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
A more realistic test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis vectors and weights with PCA.  We then use the results from this PCA to predict wavelengths for lines in the calibrations images that were left out.  The used calibration images represent the ``training set," while the left out exposures make up the ``validation set."

We chose to use 90\% of available exposures to train the PCA, and leave the remaining 10\% for validation.  Exposures were sorted into either set at random.  \lz{(Worth including?) Because only two ThAr exposures were taken each night (at the very beginning and end), leaving ThAr exposures out often left just one ThAr exposure per night.  It was therefore impossible to run a similar test using just ThAr exposures.}

We are now also testing how well PCA coefficients are interpolated to other times outside of the training set, and, consequently, how well line centers can be reconstructed for all times.  As before, differences in wavelength will also be drawn from how well line centers are measured and error in interpolating to those measured values.  This second source of error was independently quantified in the second test described in the previous section.

\subsection{Cross Tests} \label{sec:test-cross}
Similar to the training and validation tests, we can also use one calibration source to construct the PCA basis and weights and the other to test the results.  We used LFC exposures to construct a wavelength solution, and then used it to predict wavelengths for measured line centers of ThAr lines.  We can then use the assigned wavelengths for each ThAr line to assess how well we predicted the wavelengths for these lines.  Likewise, we can use ThAr exposures to predict the wavelengths of LFC lines.

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{Figures/lfcgood.pdf}
\caption{Residuals from different wavelength calibration tests.  Left: residuals of wavelength predictions for LFC lines.  In black are the results from using all LFC exposures to construct a wavelength solution, and using it to predict wavelengths for the measured line centers of each exposure.  Over-plotted in blue are the residuals from the training and validation tests.
Right: residuals of wavelength predictions for ThAr lines or using ThAr lines.  The black histogram is the same as in the left plot, but with ThAr exposures instead of LFC exposures.  In blue is the residuals from using the LFC exposures to construct a wavelength solution that is then used to predict ThAr line wavelengths.  In green is using ThAr exposures to predict LFC line wavelengths. }
\label{fig:testHists}
\end{figure}

Results from these tests are shown in Figure \ref{fig:testHists}.  The results of the self tests for LFC (left) and ThAr (right) exposures are shown as black outlines.  This histogram combines residuals from all exposures.  The self test represents the limit of how well the line centers at any given time can be predicted, since here we are predicting the line centers for an exposure used in the construction of $a_{nk}$.  \lz{How to quantify error from measurement of line center}.

Overplotted in the left plot of Figure \ref{fig:testHists} are the residuals from training the wavelength solution with only 90\% of LFC exposures and predicting the wavelengths for the remaining 10\% of exposures.  The two histograms match very closely, indicating that very little error is introduced when predicting denoised x values for times not used in the construction of the wavelength solution.

The right plot in \ref{fig:testHists} shows results involving the ThAr exposures.  Similar to with the LFC, the histogram of the ThAr self tests is shown in black, and provide a limit on how well lines at different times can be predicted.  If we use the LFCs, all taken at different times, to predict the ThAr lines, we see that again the histograms match very closely.  This confirms that interpolating $a_{nk}$ w.r.t. time is introducing little error.

Overplotted in green is the result from using ThAr lines to predict LFC line wavelengths.  These residuals show a scatter comparable to other results where the wavelength value of ThAr lines are being predicted.  This implies that using the ThAr to predict wavelengths is dominated by uncertainties in the measured line centers for the ThAr.
\lz{why is it skewed left a little??}

\section{Parametric Model} \label{sec:comparisons}
Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.  For example, the \expres\ pipeline described in \lz{Petersburg+} found the $\theta_{n}$ from equation 1 to be the amplitudes of some 2D, 9th-order polynomial in $x$ and $m$ for each $n$, making the equivalent to equation 2 the following:
\begin{equation}
\lambda(x,y,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
\quad ,
\end{equation}
where the $c_{nij}$ are coefficients unique to exposure $n$.  These coefficients are then smoothed by a third-order polynomial over time.  This same polynomial was used to construct wavelengths for non-calibration exposures.
Coefficients, $c_{nij}$, were found that minimize an objective $Q$ that is something like the L2-norm:
\begin{equation}
Q = ||\lambda(x,y,m,n) - \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j||_2^2
\quad .
\end{equation}

Even in this simple context, where every calibration image $n$ is treated as its own unique flower, there are improvements to be made.
For one, the sums should not be from 0 to 9, but instead
\begin{equation}
\sum_{i=0}^9\sum_{j=0}^{9-i}
\quad ,
\end{equation}
because that is the definition of 9th order.  We find that fitting to 8th order is sufficient, faster, and leaves less residual pattern than the original 9th degree fit.

For another, the objective function could be made soft to permit catastrophic outliers without destroying the fit. (\lz{we did not do this}).  We might recommend the iteratively reweighted least squares (IRLS).  This would make the fitting more robust.

For yet another, there are rescaling issues for the products $x^i\,m^j$ to protect the fitting from near-singularities or bad conditioning of the linear-algebra operators.

\subsection{Comparisons}
To test each method, we took an LFC exposure and separated the even and odd lines.  We then used the even lines to predict wavelengths for the odd lines and vice versa.

For example, for the parametric model, we fit a 2D, 8th-degree polynomial to the odd lines.  We then evaluated this polynomial at the positions of the even lines to assign them wavelengths.  We then reversed the process by fitting a polynomial to the even lines and evaluating that polynomial at the positions of the odd lines.  Differences between the polynomial assigned wavelengths and the assigned wavelengths of each line are shown in Figure \ref{fig:dsnMFit}.

We run the test for both the original method described in \lz{Petersburg+}, but with an 8th degree polynomial rather than 9th, and changing the polynomial for the revised description of a 2D polynomial given in equation 5.  In the second case, the polynomial is also fit using a design matrix rather than by minimizing an objective (\lz{What's the fancier way of saying this}).  For the non-parametric model, we repeat the test using a cubic spline interpolator in place of a polynomial fit.
